% There is currently a problem with spacing somewhere so that Table of
% Contents, List of Tables, and List of Figures have the wrong amount
% of space.  Others are OK though...
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

My greatest thanks goes to my advisor, Mike Whitlock. Mike has been a fantastic mentor and provided me with endless support both scientifically and academically. His thoughtful insight has been critical in the completion of the work presented herein. Mike's ability to not only teach in depth the intricacies of all aspects of population genetics, but to also enlighten me on the most tangential facts of life has made working under his supervision a great pleasure and a truly educational experience. I also thank the members of my advisory committee: Amy Angert, Loren Rieseberg, and Sally Aitken. Their helpful feedback, discussions, and support greatly improved my research.

Working amongst other members of the Otto-Whitlock labs and the UBC biodiversity research group not only made for a fantastic work environment, but also provided motivational support and vast resources of knowledge and skills. I am honored to have learned so much from everyone here and to have gained many lifelong friends and colleagues. In particular, I would like to thank Rich FitzJohn for teaching me many of the intricacies of the R language, as well as the value of good graphic design. Jeremy Draghi, who taught me C and C++, was always a great source of calm sensibility when I felt lost. Florence D\'ebarre was an outstanding mentor who imparted deep understanding of mathematical biology in the most lighthearted manners. I owe thanks to many people, both near and far, friends and family, for support at the best and worst of times -- Katie Lotterhos, Sam Yeaman, Holly Kindsvater, Nathaniel Sharp, Matt Pennell, Michael Scott, Matt Osmond, Jasmine Ono, Liz Kleynhans, Thor Veen, Carl Rothfels, Aleeza Gerstein, Kay Hodgins, Julie Lee-Yaw, Seth Rudman, Amanda Moreira, Chett Hopkins, Cathy Rushworth, Tim Vines, and Peter Fields.

Despite the absence of a chapter on local adaptation in lodgepole pine, a large portion of time during my PhD was spent on forthcoming work which would not be possible without the help of numerous people. Sally Aitken was an expert on all things trees and provided connections and support, as well as a great field vehicle. Nick Ukrainetz and Vicky Berger with the BC Ministry of Forests provided me with all details and data on the Illingworth provenance trials. I had two amazing field assistant, Evan Cronmiller and Warren Neuvonen, who were both hard-working and fun. They always managed to keep a smile on my face despite the long, hot, bug-, and pollen-filled days spent in the forests of British Columbia. Anne Berland provided excellent experience and navigational skills for visiting the Illingworth provenance trials. And Stilianos Louca, who without having crashed his car in the Yukon, I would not have succeeded in obtaining additional northern Canadian samples of lodgepole pine. Katie Lotterhos, Chett Hopkins, Roxy, Florence D\'ebarre, Holly Kindsvater, and Jeremy Draghi all also contributed their own time to help me at various times in the field. Kristin Nurkowski was my trusted expert in the wet lab, seeing me through troubleshooting DNA extractions and devising solutions to all of the difficulties that arise in the lab. I have a newfound appreciation for bunnies and a clean centrifuge thanks to Kristin. I also thank all of the members of the AdapTree team and Sally Aitken's lab.

Compute Canada and Westgrid were essential resources without which this research would not have been possible. I thank Fr\'ed\'eric Guillaume for all of his help explaining and tinkering in \textsc{nemo}.

This work was financially supported by a BRITE Fellowship, UBC Affiliated Fellowships, a Zoology Graduate Fellowship, and an NSERC Discovery Grant. 
%%% Local Variables:
%%% TeX-master: "thesis.tex"
%%% TeX-PDF-mode: t
%%% End:
