\chapter{Supplementary Information to \Chapref{expansionload}}

\section{Papers and data used in our study}
\begin{enumerate}

\item Baker A, Pearson T, Price EP, Dale J, Keim P, Hornstra H, Greenhill A, Padilla G, Warner J (2011) Molecular phylogeny of \emph{Burkholderia pseudomallei} from a remote region of Papua New Guinea. \emph{PLoS One}, 6(3), e18343. \\ doi:10.1371/journal.pone.0018343

\item Beatty GE, Provan J (2011) Comparative phylogeography of two related plant species with overlapping ranges in Europe, and the potential effects of climate change on their intraspecific genetic diversity. \emph{BMC Evolutionary Biology}, 11:29, doi:10.1186/1471-2148-11-29

\item Bilgmann K, Moeller LM, Harcourt RG, Kemper CM, Beheregaray LB (2011) The use of carcasses for the analysis of cetacean population genetic structure: A comparative study in two dolphin species. \emph{PLoS One}, 6(5), e20103. \\ doi:10.1371/journal.pone.0020103

\item Bol SM, Moerland PD, Limou S, van Remmerden Y, Coulonges C, van Manen D, Herbeck JT, Fellay J, Sieberer M, Sietzema JG, van't Slot R, Martinson J, Zagury J-F, Schuitemaker H, van't Wout AB (2011) Genome-wide association study identifies single nucleotide polymorphism in DYRK1A associated with replication of HIV-1 in monocyte-derived macrophages. \emph{PLoS One}, 6(2), e17190. doi:10.1371/journal.pone.0017190

\item Bonilla C, Hooker S, Mason T, Bock CH, Kittles RA (2011) Prostate cancer susceptibility loci identified on chromosome 12 in African Americans.  \emph{PLoS One}, 6(2), e16044. doi:10.1371/journal.pone.0016044

\item Caniato FF, Guimaraes CT, Hamblin M, Billot C, Rami J-F, Hufnagel B, Kochian LV, Liu J, Garcia AAF, Hash CT, Ramu P, Mitchell S, Kresovich S, Oliveira AC, de Avellar G, Borem A, Glaszmann J-C, Schaffert RE, Magalhaes JV (2011) The relationship between population structure and aluminum tolerance in cultivated sorghum. \emph{PLoS One}, 6(6), e20830. doi:10.1371/journal.pone.0020830

\item Colbeck GJ, Turgeon J, Sirois P, Dodson JJ (2011) Historical introgression and the role of selective vs. neutral processes in structuring nuclear genetic variation (AFLP) in a circumpolar marine fish, the capelin (\emph{Mallotus villosus}). \emph{Molecular Ecology}, 20, 1976-1987.

\item Colbeck GJ, Turgeon J, Sirois P, Dodson JJ (2011) Data from: Historical introgression and the role of selective vs. neutral processes in structuring nuclear genetic variation (AFLP) in a circumpolar marine fish, the capelin (\emph{Mallotus villosus}). Dryad Digital Repository. doi:10.5061/dryad.8616

\item Combosch DJ, Vollmer SV (2011) Population genetics of an ecosystem-defining reef coral \emph{Pocillopora damicornis }in the tropical eastern Pacific. \emph{PLoS One}, 6(8), e21200. doi:10.1371/journal.pone.0021200

\item Cox K, Broeck AV, Calster HV, Mergeay J (2011) Temperature-related natural selection in a wind-pollinated tree across regional and continental scales. \emph{Molecular Ecology}, 20, 27224-2738.

\item Cox K, Vanden Broeck A, Van Calster H, Mergeay J (2011) Data from: Temperature-related natural selection in a wind-pollinated tree across regional and continental scales. Dryad Digital Repository. doi:10.5061/dryad.9067

\item Cullingham CI, Cooke JEK, Dang S, Davis CS, Cooke BJ, Coltman DW (2011) Mountain pine beetle host-range expansion threatens the boreal forest. \emph{Molecular Ecology}, 20, 2157-2171.

\item Cullingham CI, Cooke JEK, Dang S, Davis CS, Cooke BJ, Coltman DW (2011) Data from: Mountain pine beetle host-range expansion threatens the boreal forest. Dryad Digital Repository. doi:10.5061/dryad.8677

\item Didelot X, Bowden R, Street T, Golubchik T, Spencer C, McVean G, Sangal V, Anjum MF, Achtman M, Falush D, Donnelly P (2011) Recombination and population structure in \emph{Salmonella enterica}. \emph{PLoS Genetics}, 7(7), e1002191. \\ doi:10.1371/journal.pgen.1002191

\item Erdner DL, Richlen M, McCauley LAR, Anderson DM (2011) Diversity and dynamics of a widespread bloom of the toxic dinoflagellate \emph{Alexandrium fundyense}. \emph{PLoS One}, 6(7), e22965. doi:10.1371/journal.pone.0022965

\item Gomaa NH, Montesinos-Navarro A, Alonso-Blanco C, Pic{\'o} FX (2011) Temporal variation in genetic diversity and effective population size of Mediterranean and subalpine \emph{Arabidopsis thaliana} populations. \emph{Molecular Ecology}, 20, 3540-3554.

\item Gonthier P, Garbelotto M (2011) Amplified fragment length polymorphism and sequence analyses reveal massive gene introgression from the European fungal pathogen \emph{Heterobasidion annosum} into its introduced congener \emph{H. irregulare}. \emph{Molecular Ecology}, 20, 2756-2770.

\item Groot AT, Classen A, Inglis O, Blanco CA, L{\'o}pez Jr J, Vargas AT, Schal C, Heckel DG, Sch{\"o}fl G (2011) Genetic differentiation across North America in the generalist moth \emph{Heliothis virescens} and the specialist \emph{H. subflexa}. \emph{Molecular Ecology}, 20, 2676-2692.

\item Groot A, Classen A, Staudacher H, Schal C, Heckel D (2010) Data from: Phenotypic plasticity in sexual communication signal of a noctuid moth. Dryad Digital Repository. doi:10.5061/dryad.1956

\item Guillemaud T, Blin A, Simon S, Morel K, Franck P (2011) Weak spatial and temporal population genetic structure in the rosy apple aphid, \emph{Dysaphis plantaginea}, in French apple orchards. \emph{PLoS One}, 6(6), \\ e21263. doi:10.1371/journal.pone.0021263

\item Guivier E, Galan M, Chaval Y, Xuéreb A, Salvador AR, Poulle M-L, Voutilainen L, Henttonen H, Charbonnel N, Cosson JF (2011) Landscape genetics highlights the role of bank vole metapopulation dynamics in the epidemiology of Puumala hantavirus. \emph{Molecular Ecology}, 20, 3569-3583.

\item Guivier E, Galan M, Chaval Y, Xu{\'e}reb A, Ribas Salvador A, Poulle M, Voutilainen L, Henttonen H, Charbonnel N, Cosson JF (2011) Data from: Landscape genetics highlights the role of bank vole metapopulation dynamics in the epidemiology of Puumala hantavirus. Dryad Digital Repository. \\ doi:10.5061/dryad.c6qj0

\item Gunn BF, Baudouin L, Olsen KM (2011) Independent origins of cultivated coconut (\emph{Cocos nucifera L.}) in the old world tropics. \emph{PLoS One}, 6(6), e21143. \\ doi:10.1371/journal.pone.0021143

\item Hao C, Wang L, Ge H, Dong Y, Zhang X (2011) Genetic diversity and linkage disequilibrium in Chinese bread wheat (\emph{Triticum aestivum L.}) revealed by SSR markers. \emph{PLoS One}, 6(2), e17279. doi:10.1371/journal.pone.0017279

\item Hermansen JS, S{\ae}ther SA, Elgvin TO, Borge T, Hjelle E, S{\ae}tre G-P (2011) Hybrid speciation in sparrows I: phenotypic intermediacy, genetic admixture and barriers to gene flow. \emph{Molecular Ecology}, 20, 3812-3822.

\item Hermansen JS, S{\ae}ther SA, Borge T, Elgvin TO, Hjelle E, S{\ae}tre G (2011) Data from: Hybrid speciation in sparrows I: phenotypic intermediacy, genetic admixture and barriers to gene flow. Dryad Digital Repository. doi:10.5061/dryad.k7vh9

\item Hoffman JI, Grant SM, Forcada J, Phillips CD (2011) Bayesian inference of a historical bottleneck in a heavily exploited marine mammal. \emph{Molecular Ecology}, 20, 3989-4008.

\item Hoffman JI, Grant SM, Forcada J, Phillips CD (2011) Data from: Bayesian inference of a historical bottleneck in a heavily exploited marine mammal. Dryad Digital Repository. doi:10.5061/dryad.0kj5n

\item Jacobs MMJ, Smulders MJM, van den Berg RG, Vosman B (2011) What's in a name; Genetic structure in \emph{Solanum} section \emph{Petota} studied using population-genetic tools. \emph{BMC Evolutionary Biology}, 11:42, doi:10.1186/1471-2148-11-42

\item Kanno Y, Vokoun JC, Letcher BJ (2011) Fine-scale population structure and riverscape genetics of brook trout (\emph{Salvelinus fontinalis}) distributed continuously along headwater channel networks. \emph{Molecular Ecology}, 20, 3711-3729.

\item Kanno Y, Vokoun JC, Letcher BH (2011) Data from: Fine-scale population structure and riverscape genetics of brook trout (\emph{Salvelinus fontinalis}) distributed continuously along headwater channel networks. Dryad Digital Repository. doi:10.5061/dryad.5f8s2

\item Koblm{\"u}ller S, Salzburger W, Oberm{\"u}ller B, Eigner E, Sturmbauer C, Sefc KM (2011) Separated by sand, fused by dropping water: habitat barriers and fluctuating water levels steer the evolution of rock-dwelling cichlid populations in Lake Tanganyika. \emph{Molecular Ecology}, 20, 2272-2290.

\item Koblm{\"u}ller S, Salzburger W, Oberm{\"u}ller B, Eigner E, Sturmbauer C, Sefc KM (2011) Data from: Separated by sand, fused by dropping water: habitat barriers and fluctuating water levels steer the evolution of rock-dwelling cichlid populations in Lake Tanganyika. Dryad Digital Repository. doi:10.5061/dryad.8832

\item Komiya T, Fujita S, Watanabe K (2011) A novel resource polymorphism in fish, driven by differential bottom environments: An example from an ancient lake in Japan. \emph{PLoS One}, 6(2), e17430. doi:10.1371/journal.pone.0017430

\item Kunte K, Shea C, Aardema ML, Scriber JM, Juenger TE, Gilbert LE, Kronforst MR (2011) Sex chromosome mosaicism and hybrid speciation among tiger swallowtail butterflies. \emph{PLoS Genetics}, 7(9), e1002274. doi:10.1371/journal.pgen.1002274

\item Langen K, Schwarzer J, Kullmann H, Bakker TCM, Thuenken T (2011) Microsatellite support for active inbreeding in a cichlid fish. \emph{PLoS One}, 6(9), e24689. doi:10.1371/journal.pone.0024689

\item Latch EK, Boarman WI, Walde A, Fleischer RC (2011) Fine-scale analysis reveals cryptic landscape genetic structure in desert tortoises. \emph{PLoS One}, 6(11), e27794. doi:10.1371/journal.pone.0027794

\item Lawton RJ, Messmer V, Pratchett MS, Bay LK (2011) High gene flow across large geographic scales reduces extinction risk for a highly specialized coral feeding butterflyfish. \emph{Molecular Ecology}, 20, 3584-3598.

\item Lawton RJ, Messmer V, Pratchett MS, Bay LK (2011) Data from: High gene flow across large geographic scales reduces extinction risk for a highly specialised coral feeding butterflyfish. Dryad Digital Repository. doi:10.5061/dryad.4m80r

\item Leache AD (2011) Multi-locus estimates of population structure and migration in a fence lizard hybrid zone. \emph{PLoS One}, 6(9), e25827. \\ doi:10.1371/journal.pone.0025827

\item Lee C-R, Mitchell-Olds T (2011) Quantifying effects of environmental and geographical factors on patterns of genetic differentiation. \emph{Molecular Ecology}, 20, 4631-4642.

\item Lee C, Mitchell-Olds T (2011) Data from: Quantifying effects of environmental and geographical factors on patterns of genetic differentiation. Dryad Digital Repository. doi:10.5061/dryad.6rs51

\item Lee S, Jia Y, Jia M, Gealy DR, Olsen KM, Caicedo AL (2011) Molecular evolution of the rice blast resistance gene Pi-ta in invasive weedy rice in the USA. \emph{PLoS One}, 6(10), e26260. doi:10.1371/journal.pone.0026260

\item Leite TKM, Fonseca RMC, de Franca NM, Parra EJ, Pereira RW (2011) Genomic ancestry, self-reported "color'' and quantitative measures of skin pigmentation in Brazilian admixed siblings. \emph{PLoS One}, 6(11), e27162. \\ doi:10.1371/journal.pone.0027162

\item Litvintseva AP, Carbone I, Rossouw J, Thakur R, Govender NP, Mitchell TG (2011) Evidence that the human pathogenic fungus \emph{Cryptococcus neoformans var. grubii} may have evolved in Africa. \emph{PLoS One}, 6(5), e19688. \\ doi:10.1371/journal.pone.0019688

\item Lombaert E, Guillemaud T, Thomas CE, Lawson Handley LJ, Li J, Wang S, Pang H, Goryacheva I, Zakharov IA, Jousselin E, Poland RL, Migeon A, van Lenteren J, de Clercq P, Berkvens N, Jones W, Estoup A (2011) Inferring the origin of populations introduced from a genetically structured native range by approximate Bayesian computation: case study of the invasive ladybird \emph{Harmonia axyridis}. \emph{Molecular Ecology}, 20, 4654-4670.

\item Lombaert E, Guillemaud T, Thomas CE, Lawson Handley LJ, Li J, Wang S, Pang H, Goryacheva I, Zakharov IA, Jousselin E, Poland RL, Migeon A, van Lenteren J, De Clercq P, Berkvens N, Jones W, Estoup A (2011) Data from: Inferring the origin of populations introduced from a genetically structured native range by approximate Bayesian computation: case study of the invasive ladybird \emph{Harmonia axyridis}. Dryad Digital Repository. doi:10.5061/dryad.7m0b37bn

\item Ma Y, Yang M, Fan Y, Wu J, Ma Y, Xu J (2011) Population structure of the malaria vector Anopheles sinensis (\emph{Diptera: Culicidae}) in China: Two gene pools inferred by microsatellites. \emph{PLoS One}, 6(7), e22219. doi:10.1371/journal.pone.0022219

\item Magalh{\~a}es IS, Gleiser G, Labouche A-M, Bernasconi G (2011) Comparative population genetic structure in a plant-pollinator/seed predator system. \emph{Molecular Ecology}, 20, 4618-4630.

\item Magalh{\~a}es S, Blanchet E, Egas M, Olivieri I (2011) Data from: Environmental effects on the detection of adaptation. Dryad Digital Repository. \\ doi:10.5061/dryad.1kv04

\item Martin C, Herrero M, Hormaza JI (2011) Molecular characterization of apricot germplasm from an old stone collection. \emph{PLoS One}, 6(8), e23979. \\ doi:10.1371/journal.pone.0023979

\item Moorjani P, Patterson N, Hirschhorn JN, Keinan A, Hao L, Atzmon G, Burns E, Ostrer H, Price AL, Reich D (2011) The History of African gene flow into southern Europeans, Levantines, and Jews. PLoS Genetics, 7(4), e1001373. \\ doi:10.1371/journal.pgen.1001373

\item Muscarella RA, Murray KL, Ortt D, Russell AL, Fleming TH (2011) Exploring demographic, physical, and historical explanations for the genetic structure of two lineages of greater Antillean bats. \emph{PLoS One}, 6(3), e17704. \\ doi:10.1371/journal.pone.0017704

\item Nance HA, Klimley P, Galvan-Magana F, Martinez-Ortiz J, Marko PB (2011) Demographic processes underlying subtle patterns of population structure in the scalloped hammerhead shark, \emph{Sphyrna lewini}. \emph{PLoS One}, 6(7), \\ e21459. doi:10.1371/journal.pone.0021459

\item Ng J, Glor RE (2011) Genetic differentiation among populations of a Hispaniolan trunk anole that exhibit geographical variation in dewlap colour. \emph{Molecular Ecology}, 20, 4302-4317.

\item Ozer F, Gellerman H, Ashley MV (2011) Genetic impacts of \emph{Anacapa} deer mice reintroductions following rat eradication. \emph{Molecular Ecology}, 20, 3525-3539.

\item Ozer F, Gellerman H, Ashley MV (2011) Data from: Genetic impacts of \emph{Anacapa} deer mice reintroductions following rat eradication. Dryad Digital Repository. doi:10.5061/dryad.778c9

\item Pino-Yanes M, Corrales A, Basaldua S, Hernandez A, Guerra L, Villar J, Flores C (2011) North African influences and potential bias in case-control association studies in the Spanish population. \emph{PLoS One}, 6(3), e18389. \\ doi:10.1371/journal.pone.0018389

\item Qu Y, Luo X, Zhang R, Song G, Zou F, Lei F (2011) Lineage diversification and historical demography of a montane bird \emph{Garrulax elliotii} - implications for the Pleistocene evolutionary history of the eastern Himalayas. \emph{BMC Evolutionary Biology}, 11:174, doi:10.1186/1471-2148-11-174

\item Ricca M, Sz{\"o}v{\'e}nyi P, Temsch EM, Johnson MG, Shaw AJ (2011) Interploid hybridization and mating patters in the \emph{Sphagnum subsecundum} complex. \emph{Molecular Ecology}, 20, 3203-3218.

\item Ricca M, Sz{\"o}v{\'e}nyi P, Johnson MG, Shaw AJ, Temsch EM (2011) Data from: Interploidal hybridization and mating patterns in the \emph{Sphagnum subsecundum} complex. Dryad Digital Repository. doi:10.5061/dryad.6pn84

\item Ronnås C, Cassel-Lundhagen A, Battisti A, Wall{\'e}n J, Larsson S (2011) Limited emigration from an outbreak of a forest pest insect. \emph{Molecular Ecology}, 20, 4606-4617.

\item Ronnås C, Cassel-Lundhagen A, Battisti A, Wall{\'e}n J, Larsson S (2011) Data from: Limited emigration from an outbreak of a forest pest insect. Dryad Digital Repository. doi:10.5061/dryad.gg610

\item Rossetto M, Thurlby KAG, Offord CA, Allen CB, Weston PH (2011) The impact of distance and a shifting temperature gradient on genetic connectivity across a heterogeneous landscape. \emph{BMC Evolutionary Biology}, 11:126, doi:10.1186/1471-2148-11-126

\item Sanchez-Guillen RA, Wellenreuther M, Cordero-Rivera A, Hansson B (2011) Introgression and rapid species turnover in sympatric damselflies. \emph{BMC Evolutionary Biology}, 11:210, doi:10.1186/1471-2148-11-210

\item Sandrock C, Razmjou J, Vorburger C (2011) Climate effects on life cycle variation and population genetic architecture of the black bean aphid, \emph{Aphis fabae}. \emph{Molecular Ecology}, 20, 4165-4181.

\item Sandrock C, Razmjou J, Vorburger C (2011) Data from: Climate effects on life cycle variation and population genetic architecture of the black bean aphid, \emph{Aphis fabae}. Dryad Digital Repository. doi:10.5061/dryad.2d1b8

\item Shen K-N, Jamandre BW, Hsu C-C, Tzeng W-N, Durand J-D (2011) Plio-Pleistocene sea level and temperature fluctuations in the northwestern Pacific promoted speciation in the globally-distributed flathead mullet \emph{Mugil cephalus}. \emph{BMC Evolutionary Biology}, 11:83, doi:10.1186/1471-2148-11-83

\item Takahashi MK, Horner LM, Kubota T, Keller NA, Abrahamson WG (2011) Extensive clonal spread and extreme longevity in saw palmetto, a foundation clonal plant. \emph{Molecular Ecology}, 20, 3730-3742.

\item Takahashi MK, Horner LM, Kubota T, Keller NA, Abrahamson WG (2011) Data from: Extensive clonal spread and longevity of saw palmetto (\emph{Serenoa repens}) provide insight into management plans. Dryad Digital Repository. \\ doi:10.5061/dryad.6th24

\item Trognitz B, Scheldeman X, Hansel-Hohl K, Kuant A, Grebe H, Hermann M (2011) Genetic population structure of cacao plantings within a young production area in Nicaragua. \emph{PLoS One}, 6(1), e16056. doi:10.1371/journal.pone.0016056

\item Verspoor RL, Haddrill PR (2011) Genetic diversity, population structure and Wolbachia infection status in a worldwide sample of \emph{Drosophila melanogaster} and \emph{D. simulans} populations. \emph{PLoS One}, 6(10), e26318. doi:10.1371/journal.pone.0026318

\item Via M, Gignoux CR, Roth LA, Fejerman L, Galanter J, Choudhry S, Toro-Labrador G, Viera-Vera J, Oleksyk TK, Beckman K, Ziv E, Risch N, Burchard EG, Martinez-Cruzado JC (2011) History shaped the geographic distribution of genomic admixture on the island of Puerto Rico. \emph{PLoS One}, 6(1), e16513. \\ doi:10.1371/journal.pone.0016513

\item Vigouroux Y, Mariac C, De Mita S, Pham J-L Gerard B, Kapran I, Sagnard F, Deu M, Chantereau J, Ali A, Ndjeunga J, Luong V, Thuillet A-C, Saidou A-A, Bezancon G (2011) Selection for earlier flowering crop associated with climatic variations in the Sahel. \emph{PLoS One}, 6(5), e19563. doi:10.1371/journal.pone.0019563

\item Wallace LE, Culley TM, Weller SG, Sakai AK, Kuenzi A, Roy T, Wagner WL, Nepokroeff M (2011) Asymmetrical gene flow in a hybrid Zone of Hawaiian \emph{Schiedea} (Caryophyllaceae) species with contrasting mating systems. \emph{PLoS One}, 6(9), e24845. doi:10.1371/journal.pone.0024845

\item Wellenreuther M, Sanchez-Guillen RA, Cordero-Rivera A, Svensson EI, Hansson B (2011) Environmental and climatic determinants of molecular diversity and genetic population structure in a Coenagrionid damselfly. \emph{PLoS One}, 6(6), e20440. doi:10.1371/journal.pone.0020440

\item Zaffarano PL, Queloz V, Duo A, Gruenig CR (2011) Sex in the PAC: A hidden affair in dark septate endophytes? \emph{BMC Evolutionary Biology}, 11:282, doi:10.1186/1471-2148-11-282

\item Zardi GI, Nicastro KR, Canovas F, Costa JF, Serrao EA, Pearson GA (2011) Adaptive traits are maintained on steep selective gradients despite gene flow and hybridization in the intertidal zone. \emph{PLoS One}, 6(6), e19402. \\ doi:10.1371/journal.pone.0019402

\end{enumerate}







\begin{landscape}

\renewcommand{\thetable}{D\arabic{table}}
\setcounter{table}{0}

\begin{table}[h]
\tiny
\caption[Characteristics of \textsc{structure} runs in original studies]{Characteristics of \textsc{structure} runs in original studies}
\label{tab:repro_orig}
\begin{tabular}{|p{0.03\textwidth}|p{0.05\textwidth}| p{0.042\textwidth}| p{0.048\textwidth}| p{0.085\textwidth}| 
p{0.04\textwidth}| p{0.045\textwidth}| p{0.03\textwidth}| p{0.03\textwidth}| p{0.06\textwidth}| 
p{0.075\textwidth}| p{0.105\textwidth}| p{0.09\textwidth}| p{0.076\textwidth}| p{0.04\textwidth}| p{0.11\textwidth}| }
\hline
	Study & Marker type & Number of loci & Number individuals & Number pre-determined clusters & Burn-in & MCMC rep.s & Rep.s & $K$ tested & Correlated allele freq.s? & Ancestry Model & Loc. prior? & $K$ selection method & $\Delta K$ or Likelihood plot shown? & Bar plot shown? & Chosen $K$ \\ \hline
	1 & AFLP & 252 & 179 & - & 100000 & 400000 & 8 & 1-8 & Yes & Admix. & No & P \& E & Yes & Yes & 2 \\ \hline
	 &  & 243 & 179 & - & 100000 & 400000 & 8 & 1-8 & Yes & Admix. & No & P \& E & Yes & Yes & 3 \\ \hline
	2 & SSR & 309 & 9 & 11 & 200000 & 1000000 & 5 & 1-12 & Yes & Admix. & Yes & E & Yes & Yes & 2 \\ \hline
	 &  & 309 & 9 & 11 & 200000 & 1000000 & 5 & 1-12 & Yes & Admix. & Yes & E & Yes & Yes & 3 \\ \hline
	 &  & 309 & 9 & 11 & 200000 & 1000000 & 5 & 1-12 & Yes & Admix. & Yes & E & Yes & Yes & 5 \\ \hline
	3 & SSR & 9 & 258 & 9 & 50000 & 500000 & 50 & 1-10 & Yes & Admix. & N.S. & N.S. & No & Yes & 6 \\ \hline
	 &  & 8 & 111 & 9 & 50000 & 500000 & 50 & 1-10 & Yes & Admix. & N.S. & N.S. & No & Yes & 1 \\ \hline
	4 & SSR & 12 & 233 & 5 & 1000000 & 10000000 & 10 & 1-5 & Yes & Admix. & With and w/out & E & Yes & No & 2 \\ \hline
	 &  & 11 & 233 & 5 & 1000000 & 10000000 & 10 & 1-5 & Yes & Admix. & With and w/out & E & Yes & No & 2 \\ \hline
	5 & SSR & 8 & 740 & - & 100000 & 500000 & 10 & 1-10 & Yes & Admix. & N.S. & P \& E & Yes & Yes & 2 \\ \hline
	 &  & 8 & 740 & - & 100000 & 500000 & 10 & 1-10 & Yes & Admix. & N.S. & P \& E & Yes & Yes & 4 \\ \hline
	 &  & 8 & 740 & - & 100000 & 500000 & 10 & 1-10 & Yes & Admix. & N.S. & P \& E & Yes & Yes & 1 \\ \hline
	6 & SSR & 15, 18 & 538 & 1 & 100000 & 1000000 & 10 & 1-8 & Yes & Admix. & N.S. & P \& E & Yes & Yes & 2 \\ \hline
	 &  & 18 & 181 & 3 & 100000 & 1000000 & 10 & 1-8 & Yes & Admix. & No & P \& E & Yes & Yes & 5 \\ \hline
	 &  & 18 & 102 & 2 & 100000 & 1000000 & 10 & 1-8 & Yes & Admix. & No & P & Yes & N/A & 1 \\ \hline
	7 & SSR & 317 & 15 & 47 & 100000 & 100000 & N.S. & N.S. & N.S. & Admix. & Yes & P \& E & No & Yes & 3 \\ \hline
	 &  & 317 & 15 & 47 & 100000 & 100000 & N.S. & N.S. & N.S. & Admix. & No & P \& E & No & Yes & 3 \\ \hline
	8 & SSR & 14 & 751 & 6 & 500000 & 1000000 & 10 & 1-9 & Yes & Admix. & Yes & P \& E & Yes & Yes & 2 \\ \hline
	 &  & 14 & 751 & 6 & 500000 & 1000000 & 10 & 1-9 & Yes & Admix. & No & P \& E & Yes & Yes & 2 \\ \hline
	9 & AFLP & 564 & 267 & 9 & 10,000$\dagger$ & 10,000$\dagger$ & 5 & 1-9 & No$\dagger$ & Admix.$\dagger$ & No & P & No & Yes & 4 \\ \hline
	10 & SSR & 8 & 218 & 11 & 10000 & 100000 & 10 & 1-6 & Yes & Admix. & No & P \& E & No & Yes & 2 \\ \hline
	11 & SNP & 160 & 142 & 9 & 20000 & 50000 & 10 & 2-15 & Yes & Linkage & No & Nonparam. Wilcoxon test & No & Yes & 6 \\ \hline
	12 & SSR & 16 & 310 & - & 300000 & 600000 & 100 & 1-10 & Yes & Admix. & N.S. & E & Yes & Yes & 3, 4 * \\ \hline
	13 & SSR & 21 & 246 & 1 & 100000 & 1000000 & 5 & 1-9 & Yes & Admix. & N.S. & P & Yes & No & 1 \\ \hline
	14 & SSR & 10 & 1361 & - & 10000 & 100000 & 20 & 1-6 & Yes & Admix. & No & P \& E & Yes & Yes & 3 \\ \hline
	15 & SSR & 12 & 380 & 16 & 10000 & 100000 & 20 & 1-16 & Yes & Admix. & Yes & E & No & Yes & 2 \\ \hline
	16 & SSR & 18 & 586 & 9 & 100000 & 1000000 & 20 & 1-9 & Yes & Admix. & Yes & P & Yes & Yes & 2 \\ \hline
	17 & Sequence & 153 & 550 & 35 & 500000 & 100000 & 7 & 1-35 & No & Admix. & Yes & E & No & No & 2 \\ \hline
	18 & SSR & 512 & 250 & - & 50000 & 100000 & 5 & 1-11 & Default & Admix. & No & E & Yes & No & 2 \\ \hline
	19 & SSR & 8 & 587 & 18 & 5000 & 20000 & 20 & 1-18 & Default & N.S. & N.S. & E & No & No & 2 \\ \hline
	20 & AFLP & 465 & 409 & - & 1000 & 10000 & N.S. & N.S. & No & No Admix. & N.S. & N.S. & No & No & 2 \\ \hline
	21 & AFLP & 134 & 284 & 13 & 10000 & 100000 & 20 & 1-5$\ddagger$ & Default & Admix. & With and w/out & P \& E & Yes & Yes & 5 With, 3 W/out \\ \hline
	22 & SSR & 11 & 678 & 2 & 50000 & 500000 & 10 & 2** & No & Admix. & No & N/A** & N/A & Yes & $K$  not inferred ** \\ \hline
	23 & SSR & 17 & 239 & - & 1000000 & 1000000 & 3 & 2, 3 & N.S. & N.S. & N.S. & N/A*** & N/A & No*** & $K$  not inferred*** \\ \hline
\end{tabular}

\bigskip{}
{\footnotesize 
	*   Runs were divided according to G' statistic in CLUMPP to produce two different conclusions, but method was not clear enough to reproduce the analysis   \\ 
	**  \textsc{structure} was run only at $K=2$ to identify hybrid individuals.  \\ 
	***  Compared the results of $K=2$ and $3$ based on findings of a previous study, only Q values were provided.  \\ 
	$\dagger$  Reanalysis was performed with burn-in of 100,000 and MCMC repetitions of 1,000,000; not all parameter settings were stated yet were obtained from email correspondence.  \\ 
	$\ddagger$  Range of $K$ tested obtained from likelihood plot; not stated in the text.  \\ 
	N.S.  Not stated in paper  \\ 
	N.P.  Figure not provided \\ 
}
\end{table}
\end{landscape}



\renewcommand{\thetable}{D\arabic{table}}

\begin{table}[h]
\footnotesize
\caption[Characteristics of \textsc{structure} runs reanalyzed]{Characteristics of \textsc{structure} runs reanalyzed.  Re-analyses followed the same settings as the original paper, using defaults when nothing was specified.}
\label{tab:repro_new}
\begin{tabular}{|p{0.1\textwidth}|p{0.15\textwidth}|p{0.12\textwidth}|p{0.16\textwidth}|p{0.12\textwidth}|p{0.12\textwidth}|}
\hline
	Study & Chosen $K$ & $K$ matches? & $\Delta K$ or likelihood plot matches? & Bar plot matches? & Reproduced? \\ \hline \hline
	1 & 2 & Yes & Yes & Yes & Yes \\ \hline
	 & 4 & No & No & No & No \\ \hline
	2 & 2 & Yes & Yes & Yes & Yes \\ \hline
	 & 3 & Yes & Yes & Yes & Yes \\ \hline
	 & 4 & No & No & No & No \\ \hline
	3 & 2 & No & - & No & No \\ \hline
	 & 1 & Yes & - & Yes & Yes \\ \hline
	4 & 2 & Yes & Yes & - & Yes \\ \hline
	 & 2 & Yes & Yes & - & Yes \\ \hline
	5 & 2 & Yes & Yes & Yes & Yes \\ \hline
	 & 4 & Yes & Yes & Yes & Yes \\ \hline
	 & 1 & Yes & Yes & Yes & Yes \\ \hline
	6 & 2 & Yes & Yes & Yes & Yes \\ \hline
	 & 4 & No & - & - & No \\ \hline
	 & 1 & Yes & Yes & - & Yes \\ \hline
	7 & 3 & Yes & - & Yes & Yes \\ \hline
	 & 2 & No & No & No & No \\ \hline
	8 & 2 & Yes & Yes & Yes & Yes \\ \hline
	 & 2 & Yes & Yes & Yes & Yes \\ \hline
	9 & 2 & No & - & No & No \\ \hline
	10 & 2 & Yes & - & Yes & Yes \\ \hline
	11 & 2 & No & - & No & No \\ \hline
	12 & 3, 5 * & No & No & No & No* \\ \hline
	13 & 1 & Yes & Yes & - & Yes \\ \hline
	14 & 2 & No & No & No & No \\ \hline
	15 & 2 & Yes & - & Yes & Yes \\ \hline
	16 & 2 & Yes & Yes & Yes & Yes \\ \hline
	17 & 2 & Yes & - & - & Yes \\ \hline
	18 & 2 & Yes & - & - & Yes \\ \hline
	19 & 2 & Yes & - & - & Yes \\ \hline
	20 & Data was not available &  &  &  & N/A \\ \hline
	21 & Data was not available &  &  &  & N/A \\ \hline
	22 & - & - & - & - & N/A** \\ \hline
	23 & 2 & N/A & - & - & N/A*** \\ \hline
\end{tabular}
\bigskip{}
{\footnotesize 
\\ * Runs were divided according to G' statistic in CLUMPP to produce two different conclusions, but method was not clear enough to reproduce the analysis \\ 
** \textsc{structure} was run only at $K=2$ to identify hybrid individuals.  \\ 
*** Compared the results of $K=2$ and $3$ based on findings of a previous study, only Q values were provided.  \\ 
}
\end{table}
